\documentclass[12pt]{amsart}

\title{Computing with Riemann Surfaces and Abelian Functions}

\author{Chris Swierczewski}

%\date{\today}

\begin{document}

\pagenumbering{gobble}

\maketitle

Riemann surfaces naturally appear in many branches of mathematics such as in the
construction of periodic solutions to integrable non-linear partial differential
equations and in the study of moduli spaces of algebraic curves and Jacobian
varieties in algebraic geometry and number theory. Although these topics have
been studied since Riemann himself it is only recently that the necessary tools
have been developed to allow us to approach these problems from a computational
perspective.

In this dissertation I develop the software infrastructure and improve upon
existing algorithms in order to provide a general-purpose computational
framework for working with Riemann surfaces given by the desingularization and
compactification of complex plane algebraic curves. These symbolic-numerical
algorithms are implemented in Abelfunctions [1], a Sage open-source mathematics
software library [2]. Specifically, I have provided an easy to use
object-oriented interface for computing the period matrix of a Riemann surface
and subsequently constructing of its Jacobian, performing arithmetic within
divisor group including the calculation of valuation divisors of meromorphic
differentials, numerically evaluating the Abel map and Riemann constant vector,
and quickly and accurately evaluating the Riemann theta function: a special
function from which Abelian functions are created.

The tools provided in Abelfunctions are applied to solving two problems. First,
I compute finite genus solutions to the Kadomtsev-Petviashvili equation: a
nonlinear 2+1-dimensional partial differential equation describing the
propagation of 2D shallow water waves. Second, I compute determinantal
representations of smooth plane curves: complex matrices $A$, $B$, and $C$ such
that $f(x,y) = \det(Ax + By + C)$.

\vfill

\tiny

\noindent [1] C. Swierczewski {\it et al.}, {\it Abelfunctions: A library for
  computing with Abelian functions, Riemann surfaces, and algebraic curves},
2016 {\tt http://github.com/abelfunctions/abelfunctions}. \\

\noindent [2] W. A. Stein {\it et al.}, {\it Sage Mathematics Software (Version
  7.1)}, The Sage Development Team, 2016, {\tt http://www.sagemath.org}.

\end{document}